---
title: "Class 11: Structural Bioinformatics"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The PDB Database 

The [PDB] (http://www.rcsb.org/) is the main repository for biomolecular structure data. 

Here we examine the contents of the PDB:

```{r}
db<- read.csv("Data Export Summary.csv", row.names=1)
head(db)
```

How many are X-ray
```{r}
(db$Total/sum(db$Total)) *100
```

What percent are protein 
```{r}
sum(db$Proteins/sum(db$Total)) *100
```

We could also try 
"Addins"> "Paste as data.frame"
```{r}
library(datapasta)

tmp <-data.frame(stringsAsFactors=FALSE,
   Experimental.Method = c("X-Ray", "Other", "NMR", "Multi Method",
                           "Electron Microscopy", "Total"),
              Proteins = c(126880, 256, 11062, 129, 2277, 140604),
         Nucleic.Acids = c(2012, 4, 1279, 5, 31, 3331),
    ProteinComplex = c(6547, 6, 259, 2, 800, 7614),
                 Other = c(8, 13, 8, 1, 0, 30),
                 Total = c(135447, 279, 12608, 137, 3108, 151579)
)

tmp

```


```{r}

```

There are 1157 as of 2019-05-07 See:
http://www.rcsb.org/pdb/results/results.do?tabtoshow=Current&qrid=81599ADF


## Section 3 Using Bio3d
```{r}
library(bio3d)

pdb<- read.pdb("1hsg") 
pdb
```

## Q6. How many amino acid residues are there in this pdb object and what are the two nonprotein
residues? 
Theres 198 protein residues 
The two nonprotein residues are HOH (127) and MK1(1)

## Q7. What type of R object is pdb$atom? HINT: You can always use the str() function to get a useful summery of any R object.

```{r}
attributes(pdb)
```

```{r}
str(pdb$atom)
```


```{r}
pdb$atom$resid
```


Atom selection is done via the function **atom.select()**

```{r}
prot.pdb<- atom.select(pdb, "protein", value= TRUE)
write.pdb(prot.pdb, file="1hsg_protein.pdb")
```


```{r}
lig.pdb<- atom.select(pdb, "ligand", value= TRUE)
write.pdb(lig.pdb, file = "1hsg_protein.pdb")
```


Section 5


```{r}
aa <- get.seq("1ake_A")
```

```{r}
b <- blast.pdb(aa)
```

