---
title: 'Class 8: Machine Learning'
author: "Enrique Sandoval"
date: "4/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##K-means clustering 

lets start with an example of running **kmeans()** function 

```{r}
#Generate some example data for clustering 
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```

```{r}
km<- kmeans(x, centers=2, nstart= 20)
km
```




#To Do:
-Use the kmeans() function setting k to 2 and nstart=20
-Inspect/print the results

>Q. How many points are in each cluster?
  #30 points in each cluster
  
```{r}
km$size
```

>Q. What ‘component’ of your result object details
- cluster size?
#2 clusters, defined by center
- cluster assignment/membership?
#clustering vector, which points got assigned to each 
- cluster center?
#cluster mean

```{r}
km$cluster 
```

```{r}
km$centers
```

>Plot x colored by the kmeans cluster assignment and
add cluster centers as blue points

```{r}
plot(x, col=km$cluster)
points(km$centers, pch= 18, col= "blue", cex= 3)
```


#Hierarchial clustering example 

```{r}
#Distance matrix calculation 
d<- dist(x)

#Clustering 
hc<- hclust(d)
plot(hc)
```



```{r}
plot(hc)
abline(h= 6, col= "red")
```


```{r}
cutree(hc, h= 6)
```


```{r}
cutree(hc, k=2)
```


```{r}
cutree(hc, k=3)
```

#Another exmample but more real life like with overlapping 
```{r}
x <- rbind(
matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```

#Q. Use the dist(), hclust(), plot() and cutree()
functions to return 2 and 3 clusters
```{r}
hc<- hclust(dist(x))
plot(hc)
abline(h=2, col= "red")
abline(h=2.8, col= "green")
```


```{r}
gp2<- cutree(hc, k=2)
gp3<- cutree(hc, k=3)
```

```{r}
gp2
```

```{r}
gp3
```


#Q. How does this compare to your known 'col' groups?
```{r}
table(gp2)
```

```{r}
table(gp3)
```


#PCA: Principal Component Analysis 

We will use the **prcomp()**
```{r}
#Example data 
## You can also download this file from the class website!
mydata <- read.csv("https://tinyurl.com/expression-CSV",
row.names=1)
head(mydata)
```


100 genes in dataset 
```{r}
nrow(mydata)
```


```{r}
ncol(mydata)
```


```{r}
colnames(mydata)
```

#Run our PCA analysis on the trasnpose of the data
```{r}
pca<- prcomp(t(mydata), scale= TRUE)
pca
```

#a basic PC1 vs PC2 2-D plot 
```{r}
plot(pca$x[,1], pca$x[,2])
```


#Percent variance is often more informative to look at 
Calculate the percent variance in each PC
```{r}
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
pca.var.per
```


```{r}
barplot(pca.var.per, main="Scree Plot",
xlab="Principal Component", ylab="Percent Variation")
```

#More useful plot 
```{r}
## A vector of colors for wt and ko samples
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
xlab=paste0("PC1 (", pca.var.per[1], "%)"),
ylab=paste0("PC2 (", pca.var.per[2], "%)"))
```


#England Data example 
```{r}
x <- read.csv("UK_foods.csv")
head(x)
```

#food stuff as rownames 
```{r}
x <- read.csv("UK_foods.csv", row.names= 1)
head(x)
```

#row and col number 
```{r}
nrow(x)
ncol(x)
```

#Spotting Differences and Trends 
```{r}
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))
```

#Chaning beside from T to F keeps all the data together for 1 plot per country 

#PCA to the rescue 
```{r}
pca<- prcomp(t(x))
summary(pca)
```

#PC1 vs PC2
```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x))
```

#Customized colors 
```{r}

colnames<- c("orange", "red", "blue", "darkgreen")

```


#Digging Deeper
```{r}
v <- round( pca$sdev^2/sum(pca$sdev^2) * 100 )
v
```

```{r}
z <- summary(pca)
z$importance
```

```{r}
## Lets focus on PC1 as it accounts for > 90% of variance 
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```












