---
title: "Class 12: Structural Bioinformatics (Pt 2)"
author: "Enrique Sandoval"
date: "5/9/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup for Docking 

We will first prepare our HIV-Pr system for drug docking by making a protein only PDB format file(i.e. we will remove water, existing ligands, etc)

```{r}
library(bio3d)
file<- get.pdb("1hsg")
```

```{r}
pdb<- read.pdb(file)
pdb
```

 Call: read.pdb(file=file)
 
```{r}
prot<- atom.select(pdb, "protein", value = TRUE) 
write.pdb(prot, file= "1hsg_protein.pdb")
prot
```
 
```{r}
lig<- atom.select(pdb, "ligand", value= TRUE)
write.pdb(lig, file= "1hsg_ligand.pdb")
lig
```
 
 We wil load these into ADT to add hydrogen and make PDBQT files for each 
 
 Now we will run outdoc vina with the cmd
 
## Process our docking results 

```{r}
library(bio3d)
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")
```



## Normal mode analysis for flexibility prediction 
```{r}
pdb <- read.pdb("1hel")
modes <- nma( pdb )
m7 <- mktrj(modes, mode=7, file="mode_7.pdb")
```

 
 
 
 
 
 
 
 
 
 
 