---
title: "Class 18: Investigating cancer genomics datasets"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exploring the Cancer Genome Atlas

Install the packages we need 

```{r}
library(GenomicDataCommons)
library(TCGAbiolinks)
library(maftools)
```

Can we talk with the NCI

```{r}
GenomicDataCommons::status()
```

##  Querying the GDC from R

```{r}
cases_by_project <- cases() %>%
  facet("project.project_id") %>%
  aggregations()
head(cases_by_project)
```


```{r}
x <- cases_by_project$project.project_id

# Make a custom color vector for our plot
colvec <- rep("lightblue", nrow(x))
colvec[29] <- "red"

# Plot with 'log' for y axis and rotate labels with 'las'
#par(___)  
barplot(x$doc_count, names.arg=x$key, log="y", col=colvec, las=2)
```

## Part 2: Designing a personalized cancer vaccine

Read in FASTA protein file sequence for both healthy and tumor tissue

```{r}
bio3d::read.fasta("lecture18_sequences.fa")
```

```{r}
library(bio3d)
seqs <- read.fasta("lecture18_sequences.fa")
seqs
```

## Identify tumor specific 9-mer peptides
Now to find where the sequences differ 
```{r}
ide<- conserv(seqs, method= "identity")
ide
```

```{r}
mis.match.sites <- which(ide< 1)
gap.sites <- gap.inspect(seqs)$t.inds
```

```{r}
gap.sites
```


```{r}
tumor.sites<- mis.match.sites[!mis.match.sites%in%gap.sites]
tumor.sites
```

Now get the sequence with all possible 9-mers

```{r}
start.position <- tumor.sites- 8
end.position <- tumor.sites +8
```

```{r}
start.position[1]: end.position[1]
```

Get the tumor sequence

```{r}
seqs$ali[2, start.position[1]: end.position[1] ]
```







