---
title: 'Class 7: R Functions and Packages'
author: "Enrique Sandoval"
date: "4/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Functions revisited

We will source a file form online with our functions from last class

```{r}
source("http://tinyurl.com/rescale-R")
```

Try out the last day's rescale() function 

```{r}
rescale(1:10)
```

Try the rescale2()function that catches string inputs 
```{r}
x <- c(1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```

```{r}
is.na(x)
```

```{r}
is.na(y)
```

Try putting these together with an AND

```{r}
is.na(x) & is.na(y)
```


Take the sum() to find out how many TRUE values we have and thus how many NAs we had in both x and y 
```{r}
sum(is.na(x) & is.na(y))
```


Now I can make this into our first function...
```{r}
both_na <- function(x,y){
  sum(is.na(x) & is.na(y))
}
```

```{r} 
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
```

```{r}
both_na(x, y2)
```

```{r}
3==3
```

```{r}
3 !=2
```

```{r}
length(y2)
```


Now lets try both_na2() function on our different lenght input vectors 
```{r}
both_na2(x,y2)
```

```{r}

which( c(F, F, T, F, T))

#which(is.na(c(1,2,NA, 4)))
```


```{r}
x <- c(1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)

both_na3(x,y)
```

```{r}
df1
```

```{r}
df2
```

Make things simple
```{r}
x <- df1$IDs
y <- df2$IDs

x
y
```


```{r}
intersect(x,y)
```

```{r}
which(x %in%  y)
```

```{r}
  gene_intersect <- function(x, y) {
    cbind(x [x%in% y], 
          y[y%in% x])
  }
```

```{r}
gene_intersect(df1$IDs, df2$IDs)
```

```{r}
gene_intersect2(df1, df2)
```

```{r}
merge(df1, df2, by= "IDs")
```

##Grade Function
Find average grade dropping the worst homework score 
```{r}
x <- c(100, 100, 100, 100, 100, 100, 100, 90)

```

```{r}
(sum(x)- min(x)) / (length(x)-1)
```

```{r}
y<- c(100, 90, 90, 90, 90, 90, 97, 80)
```





